Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-22 paper/graph.tex
A few notation fixes.
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@674 3fc257bb-8550-0410-81b3-a5c103da83f7

Input: \\
\> Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
\> Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = NULL \\
\> Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
\> MapReduce object $y$ = vector, initially empty \\
\> User-defined parameter $\alpha$ \\
\\
while $({residual} > {tolerance})$ \\
\> 1 Compute adjustment for random jumps and zero-outdegree vertices \\
\> \> Add $x$ to $MT$ \\
$*$\> \> Convert $MT$:  Key is vertex $i$ \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$, MultiValue = $(x_i, NULL)$ if $i \in {MT}$; else MultiValue = $x_i$ \\
\> \> \> If (${nvalues} = 2$) $c_{dot} = c_{dot} + x_i$; else $c_{dot} = c_{dot} + \alpha$ \\
\> \> \> Output: Key = $i$, Value = NULL if ${nvalues} = 2$ \\
\> \> $c_{dot}$ = MPI\_Allreduce($c_{dot}$, MPI\_SUM) / $|V|$ \\
\> 2 Compute $y = A^T x$  \\
\> \> Copy $x$ to $y$ \\
\> \> Add $A$ to $y$ \\
$*$\> \> Convert $y$:  Key is row index of $A$ \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$, MultiValue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
\> \> \>  Output: Key = $j$, Value = $A_{ij} y_i$ \\
\> \> Collate $y$:  Key is column index of $A$ \\
\> \> Reduce $y$: \\
\> \> \> Input: Key = $j$, MultiValue = $A_{ij} y_i$ for nonzeroes $A_{ij}$ in column $j$ of $A$ \\
\> \> \> Output: Key = $j$, Value = $\sum_i A_{ij} y_i$ \\
\> 3 Add $c_{dot}$ to $y$; compute $y_{max}$ = max element in $y$ \\
\> \> Map $y$:  \\
\> \> \> Input: Key = $i$, Value = $y_i$ \\
\> \> \> $y_i = y_i + c_{dot}$; $y_{max} = {\rm max}(y_i,y_{max})$ \\
\> \> \> Output:  Key = $i$, Value = $y_i$ \\
\> \> $y_{max}$ = MPI\_Allreduce($y_{max}$, MPI\_MAX) \\
\> 4 Scale $y$ by $y_{max}$ \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i / y_{max}$ \\
\> 5 Compute residual \\
\> \> Add $y$ to $x$ \\
$*$\> \> Convert $x$:  Key is vertex $i$ \\
\> \> Reduce $x$:  \\
\> \> \> Input: Key = $i$, MultiValue = $x_i, y_i$ \\
\> \> \> $residual = {\rm max}(|x_i - y_i|,residual)$ \\
\> \> $residual$ = MPI\_Allreduce($residual$, MPI\_MAX) \\
\> 6 Copy $y$ to $x$ \\
\\
Output: \\
\> MapReduce object $x$ = PageRank vector: Key = $i$, Value = $x_i$ = numeric rank 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
\> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ \\
\> MapReduce object $W$ = workspace
\\
Iterate: \\
\> 1 Map $W$ using $E$ as input: Convert edges to vertices \\
    \> \> Input: Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
    \> \> Output: Key = $V_i$, Value = $E_{ij}$ \\
    \> \> Output: Key = $V_j$, Value = $E_{ij}$ \\
\> 1 Add $V$ to $W$ \\
\> 1 Collate $W$: Vertex as key \\
\> 1 Reduce $W$: Emit edges of each vertex with zone of vertex \\
    \> \> Input: Key = $V_i$, MultiValue = $E E E E ... Z$ \\
    \> \> for each $E$ in MultiValue: \\
    \> \> \> Output: Key = $E_{ij}$, Value = $Z_i$ \\
\> 2 Collate $W$: Edge as key \\
\> 2 Reduce $W$: Emit zone re-assignments \\
       \> \> Input: Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
       \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
       \> \> if $Z_i$ and $Z_j$ are different: \\
       \> \> \> Output: Key = $Z_{loser}$, Value = $Z_{winner}$ \\
\> 2 Exit: if $W$ is empty (no output by Reduce 2) \\
\> 3 Map $V$: Invert vertex/zone pairs \\
    \> \> Input: Key = $V_i$, Value = $Z_i$ \\
    \> \> if $Z_i$ is not partitioned: \\
    \> \> \> Output: Key = $Z_i$, Value = $V_i$ \\
    \> \> else: \\
    \> \> \> Output: Key = $Z_i^+$, Value = $V_i$ for a random processor \\
\> 4 Map $V$ using $W$ as input: Add zone reassignments in $W$ to $V$ \\
    \> \> if $Z_i$ is not partitioned: \\
    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
    \> \> else: \\
    \> \> \> Output: Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
\> 4 Collate $V$: Zone ID as key \\
\> 4 Reduce $V$: Emit new zone assignment of each vertex \\
       \> \> Input: Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
       \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
       \> \> partition $Z_{new}$ if number of $V >$ threshold \\
       \> \> for each $V$ in MultiValue: \\
       \> \> \> Output: Key = $V_i$, Value = $Z_{new}$ \\
\\
Output: \\
\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
\> MapReduce object $V$ = maximal independent set vertices, initially empty \\
\> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
Map $E$: Assign random values consistently to each vertex $V_i$ and $V_j$ in $E$ \\
Clone $E$: Convert KV pairs in $E$ directly to KMV pairs with multivalue = NULL \\
\\
while $N_{edges}$ in $E > 0$: \\
\> 1 Reduce $E$: Determine WINNER/LOSER vertex of each edge \\
       \> \> Input: Key = $E_{ij}$, MultiValue = NULL or LOSER \\
       \> \> if multivalue = LOSER, emit nothing \\
       \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
       \> \> Output: Key = $V_w$, Value = ($V_l$, WINNER) \\
       \> \> Output: Key = $V_l$, Value = ($V_w$, LOSER) \\
\> 1 Collate $E$: Vertex as key \\
\> 2 Reduce $E$: Find WINNER vertices \\
       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> if all $V$ are flagged with WINNER, $V_i$ is a WINNER \\
       \> \> \> for each $V_j$ in MultiValue: \\
       \> \> \> \> Output: Key = $V_j$, Value = ($V_i$, WINNER) \\
       \> \> else: \\
       \> \> \> for each $V_j$ in MultiValue: \\
       \> \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
\> 2 Collate $E$: Vertex as key \\
\> 3 Reduce $E$: Find LOSER vertices \\
       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> if any $V$ is flagged with WINNER, $V_i$ is a LOSER \\
       \> \> \> for each $V_j$ in MultiValue: \\
       \> \> \> \> Output: Key = $V_j$, Value = ($V_i$, LOSER) \\
       \> \> else: \\
       \> \> \> for each $V_j$ in MultiValue: \\
       \> \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
\> 3 Collate $E$: Vertex as key \\
\> 4 Reduce $E$: Emit WINNER vertices and LOSER edges \\
       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> if all $V$ are flagged with LOSER, $V_i$ is a WINNER \\
       \> \> \> Output: to $V$: Key = $V_i$, Value = NULL \\
       \> \> for each $V_j$ in MultiValue: \\
       \> \> \> if $V_j$ is flagged with LOSER: \\
       \> \> \> \> Output: to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
       \> \> \> else: \\
       \> \> \> \> Output: to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = NULL \\
\> 4 Collate $E$: Edge as key \\
\\
Output: \\
\> MapReduce object $V$ = maximal independent set vertices: Key = $V_i$, Value = NULL 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
\> MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
\> MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
\> \> Key = $V_i$, Value = $\infty$, except Value for $V_s$ = 0 \\
\\
while not $done$: \\
\> $done$ = 1 \\
\> Map $V$ using $E$ as input: Add edge weights in $E$ to $V$ \\
\> \> Input: Key = $E_{ij}$, Value = $w_{ij}$ \\
\> \> Output: Key = $V_i$, Value = $(V_j, w_{ij})$ \\
\> Collate $V$: Vertex is key \\
\> Reduce $V$:  Loop over edges of vertex to update distance to $V_s$ \\
\> \> Input: Key = $V_i$, MultiValue = $d_i$, edges $(V_j, w_{ij})$, candidate distances $d_j$ \\
\> \> $d_{min} = {\rm min}(d_i,d_j,d_j,\ldots)$ \\
\> \> If $d_{min} \ne d_i$: \\
\> \> \>  $done$ = 0 \\
\> \> \>  Output: Key = $V_j$, Value = $d_{min} + w_{ij}$ for each edge vertex $V_j$ \\
\> \>  Output:  Key = $V_i$, Value = $d_{min}$ \\
\> $done$ = MPI\_Allreduce($done$,MPI\_MIN) \\
\\
Output: \\
\> MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
\> \> Key = $V_i$, Value = $d_i$ 
  \end{tabbing}
 \end{minipage}}\end{center}

>>>>

Input: \\
\> Pre-aggregated MapReduce object $A$ = $|V| \times |V|$ matrix:  Key = $i$, Value = $[j, A_{ij}]$ \\
\> Pre-aggregated MapReduce object $MT$ = indices of all-zero rows of $A$:  Key = $i$, Value = NULL \\
\> Pre-aggregated MapReduce object $x$ = initial PageRank vector:  Key = $i$, Value = $1/|V|$ \\
\> MapReduce object $y$ = vector, initially empty; user-defined parameter $\alpha$ \\
\\
while $({residual} > {tolerance})$ \\
\> 1 Compute adjustment for random jumps and zero-outdegree vertices \\
\> \> Add $x$ to $MT$ \\
$*$\> \> Convert $MT$:  Key is vertex $i$ \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $i$, MultiValue = $(x_i, NULL)$ if $i \in {MT}$; else MultiValue = $x_i$ \\
\> \> \> If (${nvalues} = 2$) $c_{dot} = c_{dot} + x_i$; else $c_{dot} = c_{dot} + \alpha$ \\
\> \> \> Output: Key = $i$, Value = NULL if ${nvalues} = 2$ \\
\> \> $c_{dot}$ = MPI\_Allreduce($c_{dot}$, MPI\_SUM) / $|V|$ \\
\> 2 Compute $y = A^T x$  \\
\> \> Copy $x$ to $y$ \\
\> \> Add $A$ to $y$ \\
$*$\> \> Convert $y$:  Key is row index of $A$ \\
\> \> Reduce $y$: \\
\> \> \>  Input: Key = $i$, MultiValue = $y_i, [j, A_{ij}]$ for nonzeros $A_{ij}$ in $A$ \\
\> \> \>  Output: Key = $j$, Value = $A_{ij} y_i$ \\
\> \> Collate $y$:  Key is column index of $A$ \\
\> \> Reduce $y$: \\
\> \> \> Input: Key = $j$, MultiValue = $A_{ij} y_i$ for nonzeroes $A_{ij}$ in column $j$ of $A$ \\
\> \> \> Output: Key = $j$, Value = $\sum_i A_{ij} y_i$ \\
\> 3 Add $c_{dot}$ to $y$; compute $y_{max}$ = max element in $y$ \\
\> \> Map $y$:  \\
\> \> \> Input: Key = $i$, Value = $y_i$ \\
\> \> \> $y_i = y_i + c_{dot}$; $y_{max} = {\rm max}(y_i,y_{max})$ \\
\> \> \> Output:  Key = $i$, Value = $y_i$ \\
\> \> $y_{max}$ = MPI\_Allreduce($y_{max}$, MPI\_MAX) \\
\> 4 Scale $y$ by $y_{max}$ \\
\> \> Map $y$: \\
\> \> \> Input:  Key = $i$, Value = $y_i$ \\
\> \> \> Output:  Key = $i$, Value = $y_i / y_{max}$ \\
\> 5 Compute residual \\
\> \> Add $y$ to $x$ \\
$*$\> \> Convert $x$:  Key is vertex $i$ \\
\> \> Reduce $x$:  \\
\> \> \> Input: Key = $i$, MultiValue = $x_i, y_i$ \\
\> \> \> $residual = {\rm max}(|x_i - y_i|,residual)$ \\
\> \> $residual$ = MPI\_Allreduce($residual$, MPI\_MAX) \\
\> 6 Copy $y$ to $x$ \\
\\
Output: \\
\> MapReduce object $x$ = PageRank vector: Key = $i$, Value = $x_i$ = numeric rank 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
\> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$, Value = NULL \\
\> MapReduce object $V$ = zone assignments of vertices: Key = $V_i$, Value = $Z_i=V_i$ \\
\> MapReduce object $W$ = workspace
\\
Iterate: \\
\> 1 Map $W$ using $E$ as input: Convert edges to vertices \\
    \> \> Input: Key = $E_{ij} = (V_i, V_j)$, Value = NULL \\
    \> \> Output: Key = $V_i$, Value = $E_{ij}$ \\
    \> \> Output: Key = $V_j$, Value = $E_{ij}$ \\
\> 1 Add $V$ to $W$ \\
\> 1 Collate $W$: Vertex as key \\
\> 1 Reduce $W$: Emit edges of each vertex with zone of vertex \\
    \> \> Input: Key = $V_i$, MultiValue = $E E E E ... Z$ \\
    \> \> for each $E$ in MultiValue: \\
    \> \> \> Output: Key = $E_{ij}$, Value = $Z_i$ \\
\> 2 Collate $W$: Edge as key \\
\> 2 Reduce $W$: Emit zone re-assignments \\
       \> \> Input: Key = $E_{ij}$, MultiValue = $Z_i Z_j$ \\
       \> \> $Z_{winner}$ = min($Z_i$,$Z_j$); $Z_{loser}$ = max($Z_i$,$Z_j$) \\
       \> \> if $Z_i$ and $Z_j$ are different: \\
       \> \> \> Output: Key = $Z_{loser}$, Value = $Z_{winner}$ \\
\> 2 Exit: if $W$ is empty (no output by Reduce 2) \\
\> 3 Map $V$: Invert vertex/zone pairs \\
    \> \> Input: Key = $V_i$, Value = $Z_i$ \\
    \> \> if $Z_i$ is not partitioned: \\
    \> \> \> Output: Key = $Z_i$, Value = $V_i$ \\
    \> \> else: \\
    \> \> \> Output: Key = $Z_i^+$, Value = $V_i$ for a random processor \\
\> 4 Map $V$ using $W$ as input: Add zone reassignments in $W$ to data in $V$ \\
    \> \> if $Z_i$ is not partitioned: \\
    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
    \> \> else: \\
    \> \> \> Output: Key = $Z_i^+$, Value = $Z_{winner}$ for every processor \\
    \> \> \> Output: Key = $Z_i$, Value = $Z_{winner}$ \\
\> 4 Collate $V$: Zone ID as key \\
\> 4 Reduce $V$: Emit new zone assignment of each vertex \\
       \> \> Input: Key = $Z_i$ or $Z_i^+$, MultiValue = $V V V V ... Z Z Z ...$ \\
       \> \> $Z_{new}$ = min($Z_i$ or $Z_i^+$,Z,Z,Z,...) \\
       \> \> partition $Z_{new}$ if number of $V >$ threshold \\
       \> \> for each $V_i$ in MultiValue: \\
       \> \> \> Output: Key = $V_i$, Value = $Z_{new}$ \\
\\
Output: \\
\> MapReduce object $V$ = zone assignment of each vertex: Key = $V_i$, Value = $Z_i$ 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
\> MapReduce object $V$ = maximal independent set vertices, initially empty \\
\> MapReduce object $E$ = edges of graph: Key = $(V_i,V_j)$ with $V_i < V_j$, Value = NULL \\
Map $E$: Assign random values consistently to each vertex $V_i$ and $V_j$ in $E$ \\
Clone $E$: Convert KV pairs in $E$ directly to KMV pairs with MultiValue = NULL \\
\\
while $N_{edges}$ in $E > 0$: \\
\> 1 Reduce $E$: Determine WINNER/LOSER vertex of each edge \\
       \> \> Input: Key = $E_{ij}$, MultiValue = NULL or LOSER \\
       \> \> if MultiValue = LOSER, emit nothing \\
       \> \> $V_w$ = WINNER vertex, $V_l$ = LOSER vertex \\
       \> \> Output: Key = $V_w$, Value = ($V_l$, WINNER) \\
       \> \> Output: Key = $V_l$, Value = ($V_w$, LOSER) \\
\> 1 Collate $E$: Vertex as key \\
\> 2 Reduce $E$: Find WINNER vertices \\
       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> if all $V_j$ in MultiValue are flagged with WINNER, $V_i$ is a WINNER \\
       \> \> \> for each $V_j$ in MultiValue: \\
       \> \> \> \> Output: Key = $V_j$, Value = ($V_i$, WINNER) \\
       \> \> else: \\
       \> \> \> for each $V_j$ in MultiValue: \\
       \> \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
\> 2 Collate $E$: Vertex as key \\
\> 3 Reduce $E$: Find LOSER vertices \\
       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> if any $V_j$ in MultiValue is flagged with WINNER, $V_i$ is a LOSER \\
       \> \> \> for each $V_j$ in MultiValue: \\
       \> \> \> \> Output: Key = $V_j$, Value = ($V_i$, LOSER) \\
       \> \> else: \\
       \> \> \> for each $V_j$ in MultiValue: \\
       \> \> \> \> Output: Key = $V_j$, Value = $V_i$ \\
\> 3 Collate $E$: Vertex as key \\
\> 4 Reduce $E$: Emit WINNER vertices and LOSER edges \\
       \> \> Input: Key = $V_i$, MultiValue = $V V V V ...$ \\
       \> \> if all $V_j$ in MultiValue are flagged with LOSER, $V_i$ is a WINNER \\
       \> \> \> Output: to $V$: Key = $V_i$, Value = NULL \\
       \> \> for each $V_j$ in MultiValue: \\
       \> \> \> if $V_j$ is flagged with LOSER: \\
       \> \> \> \> Output: to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = LOSER \\
       \> \> \> else: \\
       \> \> \> \> Output: to $E$: Key = $E_{ij}$ with $V_i < V_j$, Value = NULL \\
\> 4 Collate $E$: Edge as key \\
\\
Output: \\
\> MapReduce object $V$ = maximal independent set vertices: Key = $V_i$, Value = NULL 
  \end{tabbing}
 \end{minipage}}\end{center}

@@

Input: \\
\> MapReduce object $E$ = graph edges with weights: Key = ($V_i, V_j$), Value = $w_{ij}$ \\
\> MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
\> \> Key = $V_i$, Value = $\infty$, except Value for $V_s$ = 0 \\
\\
while not $done$: \\
\> $done$ = 1 \\
\> Map $V$ using $E$ as input: Add edge weights in $E$ to data in $V$ \\
\> \> Input: Key = $E_{ij}$, Value = $w_{ij}$ \\
\> \> Output: Key = $V_i$, Value = $(V_j, w_{ij})$ \\
\> Collate $V$: Vertex is key \\
\> Reduce $V$:  Loop over edges of vertex to update distance from $V_s$ \\
\> \> Input: Key = $V_i$, MultiValue = $d_i$, edges $(V_j, w_{ij})$, candidate distances $d_j$ \\
\> \> $d_{min} = {\rm min}(d_i,d_j,d_j,\ldots)$ \\
\> \> If $d_{min} \ne d_i$: \\
\> \> \>  $done$ = 0 \\
\> \> \>  Output: Key = $V_j$, Value = $d_{min} + w_{ij}$ for each edge vertex $V_j$ \\
\> \>  Output:  Key = $V_i$, Value = $d_{min}$ \\
\> $done$ = MPI\_Allreduce($done$,MPI\_MIN) \\
\\
Output: \\
\> MapReduce object $V$ = vertices with distances from source vertex $V_s$: \\
\> \> Key = $V_i$, Value = $d_i$ 
  \end{tabbing}
 \end{minipage}}\end{center}
